---
title: "[modelimportance]{.pkg}: Evaluating model importance within a multi-model ensemble in [R]{.proglang}"
format:
  jss-pdf:
    keep-tex: true
    header-includes:
        - \usepackage{algorithm, algcompatible, setspace}
        - \renewcommand{\algorithmicrequire}{\textbf{Input:}}
        - \renewcommand{\algorithmicensure}{\textbf{Output:}}
    journal:
      type: article
      cite-shortnames: true
      suppress: [title]
      include-jss-default: false
author:
  - name: Minsu Kim
    affiliations: 
      - name: University of Massachusetts Amherst
    orcid: 0009-0008-4637-3589
    email: minsu@umass.edu
  - name: Li Shandross
    affiliations:
      - name: University of Massachusetts Amherst
    orcid: 0009-0008-1348-1954
  - name: Evan L. Ray
    affiliations: University of Massachusetts Amherst `\\`{=latex} CVS Health 
    orcid: 0000-0003-4035-0243
  - name: Nicholas G. Reich
    affiliations:
      - University of Massachusetts Amherst
    orcid: 0000-0003-3503-9899
execute:
  echo: true
  warning: false
  message: false
bibliography: references.bib
abstract: |
  Ensemble forecasts are commonly used to support decision-making and policy planning across various fields because they often offer improved accuracy and stability compared to individual models. As each model has its own unique characteristics, understanding and measuring the value each constituent model adds to the overall accuracy of the ensemble is of great interest to building effective ensembles. The [R]{.proglang} package [modelimportance]{.pkg}  provides tools to quantify how each component model contributes to the acccuracy of ensemble performance for both point and probabilistic forecasts. It supports multiple functionalities; it allows users to specify which ensemble approach to implement and which model importance metric to use. Additionally, the software offers customizable options for handling missing values. These features enable the package to serve as a versatile tool for researchers and practitioners. It helps not only in constructing an effective ensemble model across a wide range of forecasting tasks, but also in understanding the role of each model within the ensemble and gaining insights into individual models themselves. This package follows the 'hubverse' framework, which is a collection of open-source software and tools developed to promote collaborative modeling hub efforts and simplify their setup and operation.
keywords: [ensemble, forecast, prediction, model importance, Shapley value, R]
keywords-formatted: [ensemble, forecast, prediction, model importance, Shapley value, R]
---

```{r}
#| echo: false
library(hubEnsembles)
library(hubUtils)
library(hubExamples)
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(here)
library(ggplot2)
library(kableExtra)
# install.packages("devtools")
devtools::install_github("mkim425/modelimportance")
library(modelimportance)
options(tibble.width = Inf)
```


## Introduction {#sec:intro}

Ensemble forecasting is a method to produce a single, consolidated prediction by combining forecasts generated from different models. While each model's strengths are pronounced, its weaknesses are counterbalanced, which leads to an ensemble forecast that is more robust and accurate [@gneiting2005weather; @hastie01statisticallearning].
Specifically, ensembles effectively mitigate the bias and variance arising from the predictions of individual models by averaging them out, and aggregating in this way can reduce prediction errors and improve overall performance. 
Enhanced prediction accuracy and robustness enable the achievement of more reliable predictions, thereby improving decision-making. For this reason, ensemble forecasting is widely used across various domains such as weather forecasting [@Guerra_2020; @gneiting2005weather], financial modeling [@SUN2020101160; @math11041054], and infectious disease outbreak forecasting [@ray_prediction_2018; @reich_accuracy_2019; @lutz_applying_2019; @viboud_rapidd_2018]. For example, throughout the COVID-19 pandemic, the US COVID-19 Forecast Hub collected individual models developed by over 90 different research groups and built a probabilistic ensemble forecasting model for COVID-19 cases, hospitalizations, and deaths in the US based on those models’ predictions, which served as the official short-term forecasts for the US Centers for Disease Control and Prevention (CDC) [@cramer2022united].

The quality of forecasts is assessed by evaluating their error, bias, sharpness, and/or calibration using different scoring metrics.
The selection of the scoring metrics depends on the type of forecast: 
point forecasts (e.g., mean, median) and probabilistic forecasts (e.g., quantiles, samples, predictive cumulative distribution functions, probability mass function).
Commonly used assessment tools for point forecasts are the mean absolute error (MAE) and the mean squared error (MSE), which calculate the average magnitude of forecast errors. 
Proper scoring rules for probabilistic forecasts consider the uncertainty and variability in predictions and provide concise evaluations through numerical scores [@gneiting_strictly_2007].  Some examples include the weighted interval score (WIS) for the quantile-based forecasts and the continuous ranked probability score (CRPS) for the forecasts taking the form of predictive cumulative distribution functions [@bracher_evaluating_2021]. We note that CRPS is a general scoring rule that can be computed either analytically in closed form or numerically from samples, and WIS is a quantile-based approximation of CRPS.

Several [R]{.proglang} packages have been developed for this purpose. To name a few, the [fable]{.pkg} package [@Rpackage-fable] is widely used for univariate time series forecasting and includes functions for accuracy measurement. The [Metrics]{.pkg} [@Rpackage-Metrics] and [MLmetrics]{.pkg} [@Rpackage-MLmetrics] provide a wide range of performance metrics specifically designed for evaluating machine learning models. The [scoringRules]{.pkg} [@Rpackage-scoringRules] package offers a comprehensive set of proper scoring rules for evaluating probabilistic forecasts and supports both univariate and multivariate settings. 
The [scoringutils]{.pkg} [@bosse2022evaluating] package offers additional features to the functionality provided by
[scoringRules]{.pkg}, which makes it more useful for certain tasks, such as summarizing, comparing, and visualizing forecast performance. 
These packages have been valuable to evaluate individual models as independent entities, using performance metrics selected for each specific situation or problem type. However, they do not measure the individual models' contributions to the enhanced predictive accuracy when used as part of an ensemble. 
@kim2024 demonstrate that a model's individual performance does not necessarily correspond to its contribution as a component within an ensemble.
Our developed package introduces this capability.
The [modelimportance]{.pkg} package provides tools to evaluate the role of each model as an ensemble member within an ensemble model, rather than focusing on the individual predictive performance per se. 

In ensemble forecasting, certain models contribute more significantly to the overall predictions than others. Assessing the impact of each component model on ensemble predictions is methodologically similar to determining variable importance in traditional regression and machine learning models, where variable importance measures evaluate how much individual variables decrease in accuracy of the model's predictive performance or reduce the average loss. [R]{.proglang} packages such as [randomForest]{.pkg} [@Rpackage-randomForest], [caret]{.pkg} [@Rpackage-caret], [xgboost]{.pkg} [@Rpackage-xgboost], and [gbm]{.pkg} [@Rpackage-gbm] implement these functions for different types of models: random forest models, general machine learning models, extreme gradient boosting models, and generalized boosted regression models, respectively.
These packages focus on feature-level importance within a single model and do not measure the contribution of individual models within an ensemble. The [modelimportance]{.pkg} package addresses this limitation.
The tools in the [modelimportance]{.pkg} quantify how each component model helps enhance the ensemble model's predictive performance. They assign numerical scores to each model based on a selected metric that measures forecast accuracy, depending on the forecast type.


These capabilities provide unique support for hub organizers who refer to the entity responsible for launching and managing a collective modeling hub. They coordinate multiple teams to produce forecasts and integrate their predictions into an ensemble forecast [@Shandross2024], which, as mentioned earlier, is known for having better performance compared to individual models. Examples include the CDC in the US and the European Centre for Disease Prevention and Control in the EU. 
The [modelimportance]{.pkg} package can even strengthen the benefits of a multi-model ensemble by helping these organizations create more effective ensemble forecasts based on the precise evaluation of each model's contribution. Specifically, [modelimportance]{.pkg} follows 'hubverse' standards, where 'hubverse' offers a set of publicly available software and data tools developed to promote collaborative modeling hub efforts and reduce the effort required to set up and operate them [@hubverse_docs]. Adherence to the model output formats specified by the hubverse convention enables seamless integration and flexibility with other forecasting tools and systems.


The paper proceeds as follows. In Section \ref{sec:data}, we address the model output formats proposed within the hubverse framework and the structure of forecasts. Section \ref{sec:algorithms} presents two algorithms implemented in [modelimportance]{.pkg} for calculating the model importance metric: leave-one-model-out and leave-all-subsets-of-models-out. We demonstrate the various functionalities [modelimportance]{.pkg} supports in Section \ref{sec:main-function} and give some examples in Section \ref{sec:examples}. We close this paper with some concluding remarks and a discussion of possible extensions.   


## Data {#sec:data}

### Relationship and Dependencies on Hubverse {#subsec:dependence_hubverse}

The [modelimportance]{.pkg} package is designed to work with the hubverse framework and, accordingly, depends on several packages in the hubverse ecosystem, such as [hubUtils]{.pkg}, [hubEnsembles]{.pkg}, [hubEvals]{.pkg}, and [hubExamples]{.pkg}. 
The [modelimportance]{.pkg} uses a `model_out_tbl` S3 class as the model output format defined in the [hubUtils]{.pkg}, which consists of utility functions to standardize prediction files and data formats (details in Section \ref{subsec:model_output_format}).
Ensembling predictions from multiple models relies on the [hubEnsembles]{.pkg}, which offers a broadly applicable framework to construct multi-model ensembles using various ensemble methods.
Calculation of forecast accuracy using various metrics is based on the [hubEvals]{.pkg}, which internally leverage the [scoringutils]{.pkg}.
We use or modify the example datasets from the [hubExamples]{.pkg}  for testing and demonstration purposes (see Section \ref{sec:examples}). 

\textcolor{red}{Should I cite relevant papers or links for each package?}

### Model output format {#subsec:model_output_format}
Model outputs are structured in a tabular format designed specifically for predictions, which is a formal S3 object called `model_out_tbl`.
In the hubverse standard, each row represents an individual prediction for a single task, and its details are described in multiple columns through which one can identify the model IDs, task characteristics, prediction representation type, and predicted values [@Shandross2024]. 
To elaborate on the task characteristics, each prediction task means a specific forecasting problem and it is defined by a set of task ID variables, which includes a date on which forecasts are generated, the target to predict (e.g., flu-related incident deaths, cases, or hospitalizations), and the prediction horizon, which is the length of time into the future from the point when a model generate its forecast, for a specific location on a certain target date.
@tbl-example-model_output illustrates short-term forecasts of weekly incident influenza hospitalizations in the US for Massachusetts, generated by the model 'Flusight-baseline' on December 17, 2022, in the `model_out_tbl` format.
The `model_id` column lists a uniquely identified model name. All of the `reference_date`, `target`, `horizon`, `location`, and `target_end_date` columns are all referred to as the task ID variables, which together defines the task characteristics.
Note that the forecast generation date and the target date for prediction are mapped to the `reference_date` and `target_end_date` columns, respectively, and the location is represented based on the FIPS code (e.g., '25' for Massachusetts).
The time length to the `target_end_date`, which is the number of weeks ahead from the `reference_date`, is indicated in the `horizon` column. 
The prediction representation is specified as 'quantile' in the `output_type` column, and details are represented in the `output_type_id` column with seven quantiles of 0.05, 0.1, 0.25,  0.5, 0.75, 0.9, and 0.95 for each target end date.
The predicted value corresponding to each quantile is recorded in the `value` column.


```{r example-model_output}
#| echo: false
#| label: tbl-example-model_output
#| tbl-cap: "Example of the model output for incident influenza
#| hospitalizations extracted from `forecast_outputs` data in the
#| [hubExamples]{.pkg} package."

forecast_data <- hubExamples::forecast_outputs |>
  dplyr::filter(
    output_type %in% c("quantile"),
    location == "25"
  )

forecast_data |>
  filter(
    reference_date == "2022-12-17",
    horizon != 0
  ) |> 
  head(10) |>
  knitr::kable(format = "latex", booktabs = TRUE) |>
  kable_styling(latex_options = c("scale_down"))
```

@fig-example-model_output visualize the information on the prediction task provided by @tbl-example-model_output for three models. 
For each model, the quantile-based forecasts are shown for the target end dates of December 24, 2022 (horizon 1), December 31, 2022 (horizon 2), and January 07, 2023 (horizon 3), which were made on December 17, 2022 based on the historical data available as of that date. The prediction intervals defined by the lowest and highest quantiles (0.05 and 0.95) represent the uncertainty of the predictions. To give brief details in the interpretation, the Flusight-baseline model under-predicted the outcomes for the first two target dates (horizon 1 and 2), but it over-predicted the outcome for the last target date (horizon 3). Its prediction intervals are narrow compared to the other two models, which indicates that it is more confident about its predictions. 

```{r fig-example-model_output, fig.pos="t"}
#| echo: false
#| label: fig-example-model_output
#| fig-cap: "Example plot of three distributional forecasts corresponding to the model output for incident influenza hospitalizations shown in @tbl-example-model_output. Solid black dots indicate historically available data as of the forecast generation date, and open black circles indicate the eventually observed values. The blue dots represent predictive medians and the blue shaded area represents the corresponding 90% prediction interval defined by the 0.05 and 0.95 quantiles."

target_data <- hubExamples::forecast_target_ts |>
  dplyr::filter(
    date %in% unique(forecast_data$target_end_date),
    location %in% unique(forecast_data$location)
  ) |>
  rename(target_end_date = date)

forecast_data |>
  filter(
    reference_date == "2022-12-17",
    horizon != 0,
    output_type_id %in% c(0.05, 0.5, 0.95)
  ) |>
  pivot_wider(names_from = output_type_id, values_from = value) |>
  rename(lower = "0.05", upper = "0.95", value = "0.5") |>
  ggplot(aes(x = target_end_date)) +
  facet_grid(~model_id) +
  geom_point(aes(y = value, color = "medians"), size = 2) +
  geom_line(aes(y = value, color = "medians"), linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = lower, ymax = upper,
      fill = "#3388FF"
    ),
    alpha = 0.5
  ) +
  geom_point(
    data = target_data |> filter(target_end_date <= "2022-12-17"),
    aes(y = observation, group = 1, color = "obs")
  ) +
  geom_line(
    data = target_data |> filter(target_end_date <= "2022-12-17"),
    aes(y = observation, group = 1, color = "obs")
  ) +
  geom_point(
    data = target_data |> filter(target_end_date > "2022-12-17"),
    aes(y = observation, group = 1, color = "truth"),
    shape = 1, alpha = 1
  ) +
  geom_line(
    data = target_data |> filter(target_end_date > "2022-12-17"),
    aes(y = observation, group = 1, color = "truth"),
    alpha = 0.75
  ) +
  # coord_cartesian(ylim = c(0, 500)) +
  scale_x_date(breaks = target_data$target_end_date, date_labels = "%Y-%m-%d") +
  labs(
    y = "Weekly Hospitalization",
    # title = "Forecasts of incident deaths in Massachusetts on November 28, 2021",
    x = "Date"
  ) +
  scale_color_manual(
    name = "",
    values = c(
      "medians" = "DodgerBlue",
      "obs" = "Black",
      "truth" = "Black"
    ),
    labels = c(
      "Forecast (Predictive median)",
      "Observed data before forecasting",
      "Eventually observed value"
    )
  ) +
  scale_fill_manual("",
    values = "#3388FF",
    labels = "90% Prediction interval"
  ) +
  theme(
          axis.title.x = element_text(size = 8), 
    axis.title.y = element_text(size = 8),
    axis.text.x = element_text(size = 6, angle = 90),
    axis.text.y = element_text(size = 8),
    strip.text = element_text(size = 8),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.direction = "vertical",
    legend.box = "horizontal"
  ) 
```





### Forecast data representation {#subsec:model_output}

Generally, quantitative forecasts can be categorized into either point forecasts or probabilistic forecasts. 
For a specific prediction task, point forecasts, represented by a single predicted value, provide a clear and concise prediction, making them easy to interpret and communicate. 
Probabilistic forecasts, on the other hand, provide a probability distribution over possible future values, which inherently involves uncertainty. They are represented in various ways, such as probability mass functions (pmf), cumulative distribution functions (cdf), samples, or probability quantiles (or intervals). 

The `output_type` and `output_type_id` columns in the model output format, as defined by the hubverse convention, specify the forecast structure. 
Only one `output_type` is allowed, and it must be one of the 'mean', 'median', 'quantile', or 'pmf' in the [modelimportance]{.pkg} package: 'mean' or 'median' for point forecasts and 'quantile' or 'pmf' for probabilistic forecasts. 
As aforementioned, `output_type_id` column  identifies addtional detailed information, such as specific quantile levels (e.g., "0.1", "0.25", "0.5", "0.75", and "0.9") for the 'quantile' output type and categorical values (e.g., "low", "moderate", "high", and "very high") for the 'pmf' output type. The predicted values for `pmf` are constrained to be between 0 and 1, indicating the probability at each categorical level, while they are unbounded numeric otherwise.
Different output types correspond to different scoring rules for evaluating a model's prediction performance. @tbl-pair-output-scoringrule presents the output types and their associated scoring rules supported by the [modelimportance]{.pkg} package.


```{r}
#| echo: false
#| label: tbl-pair-output-scoringrule
#| tbl-cap: "Pairs of output types and their associated scoring rules for evaluating prediction performance."

data.frame(
  "Output Type" = c("mean", "median", "quantile", "pmf"),
  "Scoring Rule" = c("MSE", "MAE", "WIS", "Log Score"),
  Description = c(
    "Evaluate using the mean squared error (MSE)",
    "Evaluate using the mean absolute error (MAE)",
    "Evaluate using the weighted interval score (WIS)",
    "Evaluate using the logarithm of the probability assigned to the true outcome (LogScore)"
  ),
  check.names = FALSE
) |>
  knitr::kable(format = "latex", booktabs = TRUE) |>
  column_spec(3, width = "10cm")
```

### Oracle output data {#subsec:oracle_output_data}
The `oracle_output_data` is a data frame that contains the ground truth values for the variables used to define modeling targets. 
This data must follow the oracle output format defined in the hubverse standard, which includes independent task ID columns (e.g., `location`, `target_date`), the `output_type` column specifying the output type of the predictions and an `oracle_value` column for the observed values. 
As in the forecast data, if the `output_type` is either `"quantile"` or `"pmf"`, the `output_type_id` column is often required to provide further identifying information. 


The `model_out_tbl` and `oracle_output_data` must have the same task ID columns and `output_type`, including `output_type_id` if necessary, which are used to match the predictions with the ground truth data. 

## Algorithms {#sec:algorithms}

This section provides a brief description of the leave one model out (LOMO) and leave all subsets of models out (LASOMO) algorithms, which are used to compute the model importance score. (Details can be found in @kim2024) 

LOMO involves creating an ensemble by excluding one component model from the entire set of models. 
Let ${\cal A}$ be a set of $n$ models and $F^i$ be a forecast produced by model $i$, where $i = 1,2, \dots, n.$
Each ensemble excludes exactly one model while including all the others. Denoting by $F^{{\cal A}^{-i}}$ an ensemble forecast constructed without $F^i$ and by $F^{\cal A}$ the ensemble forecast built from the entire set of models, the importance score using LOMO is calculated as the difference in accuracy, as measured by a specific score, between $F^{{\cal A}^{-i}}$ and $F^{\cal A}$  (Algorithm \ref{alg:lomo}). 
For example, when evaluating model 1 within an ensemble of three models ($n=3$), LOMO creates an ensemble forecast $F^{\{2,3\}}$ using only $F^2$ and $F^3$. The performance of this reduced ensemble is then compared to the full ensemble forecast $F^{\{1,2,3\}}$, which incorporates all three models.

On the other hand, LASOMO involves ensemble constructions from all possible subsets of models. For each subset $S$ that does not contain the model $i$, $S \cup \{i\}$ plays a role of ${\cal A}$ in the LOMO; the score associated with the subset $S$ is the difference of measures between $F^S$ and $F^{S \cup \{i\}}$. Then, all scores are aggregated across all possible subsets that the model $i$ does not belong to (Algorithm \ref{alg:lasomo}).
For example, using the earlier setup of three forecast models, LASOMO considers three subsets, $\{2\}$, $\{3\}$, and $\{2, 3\}$, to calculate the importance score of model 1 (excluding all subsets that include model 1). The ensemble forecasts $F^{\{2\}}, F^{\{3\}}$, and $F^{\{2,3\}}$ are then compared to $F^{\{1,2\}}, F^{\{1,3\}}$, and $F^{\{1,2,3\}}$, respectively. The performance differences attributable to model 1's inclusion are aggregated, which results in the importance score for model 1. We note that the subsets may have different weights during the aggregating process. 
The [modelimportance]{.pkg} package offers two weighting options for subsets: one assigns equal (uniform) weights to all subsets, and the other assigns weights based on their size, similar to the concept of Shapley values. Users can choose one to evaluate the contribution of each model in a manner suited to their preferred framework.

\begin{algorithm}[bh!]
\caption{Importance score calculation for one prediction task using leave one model out (LOMO) algorithm} 
\label{alg:lomo}
 \begin{spacing}{1.2}
  \begin{algorithmic}[1]
   \REQUIRE{
    \Statex $\bullet$ Set of $n$ individual models, ${\cal A}=\{1,2,\dots, n\},$ and their forecasts $\{F^1,F^2,...,F^n\}$
    \Statex $\bullet$ Truth value, $y$
    \Statex $\bullet$ Function $g$ to build an ensemble
    \Statex $\bullet$ Scoring rule $\psi$ to evaluate forecast skills}
   \ENSURE importance metric of model $i$, $\phi^{i,\text{lomo}}$
    \STATE{Create an ensemble forecast $F^{\cal A}$ using $g$: $F^{\cal A} \gets g(\{F^1,F^2,...,F^n\})$}
    \STATE{Evaluate $F^{\cal A}$ using $\psi$: $\psi(F^{\cal A},y).$}
    \FOR{each $i, (i = 1,2,..., n)$}
      \STATE {$F^{{\cal A}^{-i}} \gets g(\{F^j|\, j\ne i, j\in {\cal A}\})$}
      \STATE{Compute $\psi(F^{{\cal A}^{-i}},y).$}
      \STATE $\phi^{i,\text{lomo}} \gets   \psi(F^{{\cal A}^{-i}},y) - \psi(F^{\cal A},y)$
    \ENDFOR
  \end{algorithmic}
 \end{spacing}
\end{algorithm}


Algorithms \ref{alg:lomo} and \ref{alg:lasomo} outline the steps to implement LOMO and LASOMO for a single prediction task, respectively.

\begin{algorithm}[bh!]
\caption{Importance score calculation for one prediction task using leave all subsets of models out (LASOMO) algorithm} 
\label{alg:lasomo}
 \begin{spacing}{1.2}
  \begin{algorithmic}[1]
   \REQUIRE{
    \Statex $\bullet$ Set of $n$ individual models, ${\cal A}=\{1,2,\dots, n\},$ and their forecasts $\{F^1,F^2,...,F^n\}$
    \Statex $\bullet$ Truth value, $y$
    \Statex $\bullet$ Function $g$ to build an ensemble
    \Statex $\bullet$ Scoring rule $\psi$ to evaluate forecast skills}
   \ENSURE Importance metric of model $i$, $\phi^{i,\text{lasomo}}$
   \FOR{$i = 1 \text{ to } n$}
     \STATE $\phi^{i,\text{lasomo}} \gets 0$
     \STATE Make a list of non-empty subsets of $\cal A$ that does not contain $i$: $S_1, S_2, \cdots, S_{2^{n-1}-1}$
     \FOR{$j = 1 \text{ to } 2^{n-1}-1$}
       \STATE{Assign a weight to the subset $S_j$: }
         \IF{all subsets have uniform weights}
           \STATE $\gamma_{S_j} \gets \displaystyle\frac{1}{2^{n-1}-1}$
         \ELSE{ (\textit{subset's weight depends on its size})}
           \STATE {$\gamma_{S_j} \gets \displaystyle\frac{1}{(n-1)\binom{n-1}{|S_j|}}$}
         \ENDIF
       \STATE{$F^{S_j}\gets g(\{F^j|\, j\in S_j\})$}
       \STATE{$F^{{S_j}\cup \{i\}}\gets g(\{F^i, F^j|\, j\in S_j\})$}
       \STATE{Compute $\psi(F^{S_j},y)$ and
       $\psi(F^{{S_j}\cup\{i\}},y)$}
       \STATE {$\phi^{i,\text{lasomo}} \gets \phi^{i,\text{lasomo}} + \gamma_{S_j}\times\Big[ \psi(F^{S_j},y) - \psi(F^{{S_j}\cup\{i\}},y) \Big]$}
     \ENDFOR
   \ENDFOR
  \end{algorithmic}
 \end{spacing}
\end{algorithm}


## Main function: Evaluating ensemble members {#sec:main-function}

In this section, we describe the usage of the function `model_importance()`, where multiple options are available to customize the evaluation framework (@tbl-arguments).

```r
R> model_importance(forecast_data, oracle_output_data, ensemble_fun, weighted,
                    training_window_length, importance_algorithm, subset_wt, 
                    na_action, ...)
```


```{r}
#| echo: false
#| label: tbl-arguments
#| tbl-cap: "Description of the arguments for the `model_importance()` function, including their purpose, possible values, and default settings."

tbl_argument <- data.frame(
  "Argument" = c(
    "\\texttt{forecast\\_data}", "\\texttt{oracle\\_output\\_data}",
    "\\texttt{ensemble\\_fun}", "\\texttt{training\\_window\\_length}",
    "\\texttt{importance\\_algorithm}", "\\texttt{subset\\_wt}",
    "\\texttt{na\\_action}", "\\texttt{...}"
  ),
  "Description" = c(
    "Forecasts",
    "Ground truth data",
    "Ensemble method",
    "Time interval of historical data used during the training process",
    "Algorithm to calculate importance",
    "Method for assigning weight to subsets when using LASOMO algorithm",
    "Method to handle missing data",
    "Optional arguments for \\texttt{simple\\_ensemble}"
  ),
  "Possible Values" = c(
    "Must be the model output format",
    "Must be the oracle output format",
    "\\texttt{simple\\_ensemble}, \\texttt{linear\\_pool}",
    "Non-negative integer",
    "\\texttt{lomo, lasomo}",
    "\\texttt{equal, perm\\_based}",
    "\\texttt{worst, average, drop}",
    "Varies"
  ),
  Default = c(
    "N/A",
    "N/A",
    "\\texttt{simple\\_ensemble}",
    "0",
    "\\texttt{lomo}",
    "\\texttt{equal}",
    "\\texttt{worst}",
    "\\texttt{agg\\_fun = mean}"
  ),
  check.names = FALSE
)


tbl_argument |>
  knitr::kable(format = "latex", booktabs = TRUE, escape = FALSE) |>
  column_spec(c(2, 3, 4), width = c("3.8cm", "3.5cm", "2.7cm")) |>
  kable_styling(latex_options = c("striped", "t"), font_size = 10.5)
```

`forecast_data` is a data frame containing predictions and should be or can be coerced to a `model_out_tbl` format, which is the standard S3 class model output format defined by the hubverse convention. 


The `ensemble_fun` argument specifies the ensemble method to be used for evaluating model importance. 
The currently supported methods are `"simple_ensemble"` and `"linear_pool"`. 
The `"simple_ensemble"` method returns the average of the predicted values from all component models per prediction task defined by task IDs, `output_type`, and `output_type_id` columns. 
The default aggregation function for this method is `"mean"`, but it can be customized by 
specifying additional arguments through `...`, such as `agg_fun="median"`.
When `"linear_pool"` is specified, ensemble model outputs are created as a linear pool of component model outputs. This method supports only an `output_type` of `"mean"`, `"quantile"`, or `"pmf"`.

The `weighted` argument is a logical value that indicates whether model weighting should be done when building an ensemble using the specified `ensemble_fun`. 
If it is set to `TRUE`, model weights are estimated based on the previous performance of each model, and these weights are used to build the ensemble. 

The `importance_algorithm` argument specifies the algorithm for model importance calculation, which can be either `"lomo"` (leave-one-model-out) and `"lasomo"` (leave all subsets of models out). The `subset_wt` argument is employed only for the `"lasomo"` algorithm. This argument has two options: `"equal"` assigns equal weight to all subsets and `"perm_based"` assigns weight averaged over all possible permutations as in
the formula of Shapley values (Algorithm \ref{alg:lasomo}). The default values of `importance_algorithm` and `subset_wt` are `"lomo"` and `"equal"`, respectively.


The `na_action` argument allows for specifying how to handle missing values in the `forecast_data`. 
Three options are available: `"worst"`, `"average"`, and `"drop"`. 
In each specific prediction task, if a model has any missing predictions, the `"worst"` option replaces those missing values with the smallest value from the other models, while the `"average"` option replaces them with the average of the other models' predictions in that task. The `"drop"` option removes missing values, which results in the exclusion of the model from the evaluation for that task.


## Implementation and Availability

This package is implemented in [R]{.proglang} and distributed via CRAN and GitHhub. 
We conducted unit tests using the [testthat]{.pkg} package [@Rpackage-testthat] to ensure the correctness of all functions, including those used internally. Additionally, we performed continuous integration testing using GitHub Actions to maintain code quality and functionality across platforms. 


## Examples {#sec:examples}

The examples in this section illustrate the use of the `model_importance()` function to evaluate the importance of component models within an ensemble, using various combinations of the arguments described in Section \ref{sec:main-function}. We use some example forecast and target data from the [hubExamples]{.pkg} package, which provides sample datasets for multiple modeling hubs in the hubverse format. 



### Evaluation using untrained ensemble in LOMO algorithm for mean forecasts  {#sec:untrained-lomo-mean}

The forecast data used here contains forecasts of weekly incident influenza hospitalizations in the US for Massachusetts (FIPS code 25) and Texas (FIPS code 48), generated on November 19, 2022. 
These forecasts are for two target end dates, November 26, 2022 (horizon 1), and  December 10, 2022 (horizon 3), and were produced by three models: 'Flusight-baseline', 'MOBS-GLEAM_FLUH', and 'PSI-DICE'. The output type is `mean` and the `output_type_id` column has `NA`s as no further specification is required for this output type. 
We have modified the example data slightly: the forecast values were rounded to the nearest integer and some forecasts have been removed to demonstrate the handling of missing values. Therefore, 'MOBS-GLEAM_FLUH''s forecast for Massachusetts on November 26, 2022, and 'PSI-DICE's forecast for Texas on December 10, 2022, are missing.


```r
R> forecast_data
``` 
\small
```{r tidy = FALSE}
#| echo: false
forecast_data <- hubExamples::forecast_outputs |>
  dplyr::filter(
    output_type %in% c("mean"),
    target_end_date %in% as.Date(c("2022-11-26", "2022-12-10"))
  ) |>
  mutate(value = round(value)) |>
  filter(
    !(model_id == "MOBS-GLEAM_FLUH" & location == "25" & target_end_date == as.Date("2022-11-26")),
    !(model_id == "PSI-DICE" & location == "48" & target_end_date == as.Date("2022-12-10"))
  )

print(forecast_data, width = Inf)
```

\normalsize

The corresponding target data contains the observed hospitalization counts for these dates and locations.

```r
R> target_data
``` 
\small
```{r}
#| echo: false
target_data <- hubExamples::forecast_target_ts |>
  dplyr::filter(
    date %in% unique(forecast_data$target_end_date),
    location %in% unique(forecast_data$location)
  ) |>
  # Rename columns to match the oracle output format
  rename(
    target_end_date = date,
    oracle_value = observation
  )

print(target_data, width = Inf)
```

Overall, the forecasts tend to have larger prediction errors for the target end date of December 10, 2022, compared to November 26, 2022, which is expected due to increased uncertainty at  longer horizons. Additionally, the forecasts for Massachusetts are relatively more accurate compared to those for Texas, which tend to have higher errors.

\normalsize

We can evaluate the importance of each model in the ensemble using the `model_importance()` function. 
The following code evaluates the importance of each model in the simple mean ensemble using the LOMO algorithm, without training the ensemble (`weighted = FALSE`). The `na_action` argument is set to `"drop"`, which represents that any missing values in the forecasts will be excluded from the evaluation. 

```r
R> model_importance(
                    forecast_data = forecast_data, 
                    oracle_output_data = target_data,
                    ensemble_fun = "simple_ensemble", 
                    weighted = FALSE,
                    importance_algorithm = "lomo", 
                    na_action = "drop"
                   )
```

This call generates both the result and informative messages, summarizing the input data, including the number of dates on which forecasts were produced and the number of models with ids as follows.

\small
```r
Forecasts from 2022-11-19 to 2022-11-19 (a total of 1 forecast date(s)).
The available model IDs are:
 	 Flusight-baseline
	 MOBS-GLEAM_FLUH
	 PSI-DICE 
(a total of 3 models)
```
\normalsize

The function output is a data frame containing model IDs and their corresponding importance scores in the `mean_importance` column, ordered from most to least importance.
The column name `mean_importance` indicates the average of task-specific importance scores.


\small
```{r}
#| echo: false
model_importance(
  forecast_data = forecast_data, oracle_output_data = target_data,
  ensemble_fun = "simple_ensemble", weighted = FALSE,
  importance_algorithm = "lomo",
  na_action = "drop"
)
```
\normalsize
The results show that the model 'Flusight-baseline' has the highest importance score, followed by 'PSI-DICE' and 'MOBS-GLEAM_FLUH'.
That is, 'Flusight-baseline' contributes the most to improving the ensemble's predictive performance, whereas MOBS-GLEAM_FLUH, which has a negative score, detracts from the ensemble’s performance.
The low importance score of 'MOBS-GLEAM_FLUH' is mainly due to a substantially larger prediction error for Texas on the target end date of December 10, 2022, compared to other models, while its missing forecast for Massachusetts for November 26, 2022, was not factored into the evaluation. This single large error significantly affected its contribution score.


Another approach to handling missing values is to use the `"worst"` option for `na_action`, which replaces missing values with the worst (i.e., minimum) score among the other models for the same task.

```r
R> model_importance(
                    forecast_data = forecast_data, 
                    oracle_output_data = target_data,
                    ensemble_fun = "simple_ensemble", 
                    weighted = FALSE,
                    importance_algorithm = "lomo", 
                    na_action = "worst"
                   )
```

\small 
```{r}
#| echo: false
model_importance(
  forecast_data = forecast_data, oracle_output_data = target_data,
  ensemble_fun = "simple_ensemble", weighted = FALSE,
  importance_algorithm = "lomo", na_action = "worst"
)
```

\normalsize

The results show that the importance scores of 'Flusight-baseline' is unchanged because it has no missing forecast. 
We observe that the importance score of 'PSI-DICE', which was previously positive, has now decreased to a negative value when compared to the evaluation using the `"drop"` option for `na_action`. Moreover, 'MOBS-GLEAM_FLUH' still ranks the lowest, but the importance score has increased. 
This change is related to the varying forecast accuracy across different tasks.
For the target end date of November 26, 2022, in Massachusetts, most forecasts are relatively accurate. Thus, even if the 'MOBS-GLEAM_FLUH' is assigned the worst value of importance score for its missing forecast, including this value in the averaging is not detrimental to the overall importance metric; rather, it is more beneficial than excluding it. In contrast, for the target end date of December 10, 2022, in Texas, the forecasts have much larger errors across the board, and assigning the worst value of importance score to the missing forecast of 'PSI-DICE' in this task has a detrimental effect on averaging importance scores. This is because the scale of the importance scores is influenced by the magnitude of the prediction errors: for tasks with small errors, the scores remain moderate, while tasks with large errors can yield importance scores of much greater magnitude.

It is also possible to impute the missing scores with intermediate values by assigning the average importance scores of other models in the same task. This strategy may offer a more balanced trade-off by mitigating the influence of the missing data without overly penalizing or overlooking them. 


```r
R> model_importance(
                    forecast_data = forecast_data, 
                    oracle_output_data = target_data,
                    ensemble_fun = "simple_ensemble", 
                    weighted = FALSE,
                    importance_algorithm = "lomo", 
                    na_action = "average"
                   )
```

\small 
```{r}
#| echo: false
model_importance(
  forecast_data = forecast_data, oracle_output_data = target_data,
  ensemble_fun = "simple_ensemble", weighted = FALSE,
  importance_algorithm = "lomo", na_action = "average"
)
```


\normalsize


In this simple example with only three models, the ranking of the importance scores remains unchanged. However, in more complex scenarios with a larger number of models, the choice of `na_action` can impact the importance scores and their interpretation. A more extensive application can be found in @kim2024.



### Evaluation using trained ensemble in LASOMO algorithm for quantile-based forecasts  {#sec:trained-lomo-qntl}

TO DO



## Summary and discussion {#sec:discussion}

Multi-model ensemble forecasts often provide better accuracy and robustness than single models, and are widely used in decision-making and policy planning across various domains. The contribution of each component model to the accuracy of the ensemble depends on its own unique characteristics. The [modelimportance]{.pkg} package enables the quantification of the value that each component model adds to the ensemble performance in different evaluation contexts. 

The primary function of the package is `model_importance()`, which returns a data frame with component models and their importance metrics. Users can choose the various ensemble methods to apply and model importance algorithm between LOMO and LASOMO. Additionally, customizable options are available for handling missing values. These features enable the package to serve as a versatile tool to aid collaborative efforts to construct an effective ensemble model across a wide range of forecasting tasks. 
We note that unit testing with continuous integration ensures the reliability of all functions and the overall quality of code across multiple platforms. 

There is a room to enhance the current version of this package. Although this package supports four different output types ('mean', 'median', 'quantile, and 'pmf'), other output types are widely used in practice. For example, 'sample' output type is commonly used in the US Flu Scenario Modeling Hub [@FluSMH]. This format includes multiple simulated values (samples) from the forecast distribution. The `output_type_id` is specified for each sample, which typically indexes the samples or indicates their source, depending on the context. The package can be extended to support this output type, which is under consideration for future releases. 
These extensions would aim to broaden the scope of applications in real-world forecasting tasks.


## Acknowledgements {.unnumbered}

We acknowledge Zhian N. Kamvar for debugging and resolving coding issues while developing the package. 

## References {.unnumbered}
