---
title: "Simple working examples"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Simple working examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  warning = FALSE,
  message = FALSE
)
options(width = 200)
library(dplyr)
library(ggplot2)
library(kableExtra)
```


This vignette demonstrates the usage of the `modelimportance` package for 
evaluating how each component model contributes to ensemble accuracy.
We provide simple working examples to help you get started with the package. 
Detailed descriptions of the model importance metrics, algorithms, key 
functions, and in-depth interpretations of the examples covered here are 
available in the [accompanying article](../articles/modelimportance-article.html) titled '`modelimportance`: Evaluating 
model importance within a multi-model ensemble in R' under *Articles*.


## Setup

We first load the necessary packages.
```{r setup}
library(hubExamples)
library(modelimportance)
library(dplyr)
library(ggplot2)
```

We also set up a table style for displaying the data in the following section.
```{r tbl-style}
library(kableExtra)
tbl_style <- function(x, ...) {
  knitr::kable(x, format = "html") |>
    kableExtra::kable_styling(
      font_size = 12,
      bootstrap_options = c("striped", "hover", "condensed", "responsive"),
      full_width = TRUE
    )
}
```

We use some example forecast and target data from the `hubExamples` package, 
which provides sample datasets for multiple modeling hubs in the hubverse format. 


## Example data

The forecast data used here contains forecasts of weekly incident influenza  hospitalizations in the US for Massachusetts (FIPS code 25) and Texas (FIPS code 48), generated on November 19, 2022. 
These forecasts are made for two target end dates, November 26, 2022 (horizon 1), and December 10, 2022 (horizon 3), and were produced by three models: 'Flusight-baseline', 'MOBS-GLEAM_FLUH', and 'PSI-DICE'. 
The output type is `median` and the `output_type_id` column has `NA`s as no further specification is required for this output type. 
We have modified the example data slightly by removing some forecasts to demonstrate the handling of missing values. 
Therefore, MOBS-GLEAM_FLUH's forecast for Massachusetts on November 26, 2022, and PSI-DICE's forecast for Texas on December 10, 2022, are missing.


```{r forecast_data}
# Load the forecast data from hubExamples
forecast_data_raw <- hubExamples::forecast_outputs |>
  dplyr::filter(
    .data$output_type %in% c("median"),
    .data$target_end_date %in% as.Date(c("2022-11-26", "2022-12-10"))
  )

# Specify forecasts to remove: MOBS-GLEAM_FLUH for location 25 on 2022-11-26,
# PSI-DICE for location 48 on 2022-12-10
forecast_to_remove <- tibble(
  model_id = c("MOBS-GLEAM_FLUH", "PSI-DICE"),
  location = c("25", "48"),
  target_end_date = as.Date(c("2022-11-26", "2022-12-10"))
)

# Filter out the specified forecasts from the original data
forecast_data <- forecast_data_raw |>
  anti_join(forecast_to_remove, by = c("model_id", "location", "target_end_date"))

# Display the forecast data
tbl_style(forecast_data)
```

The corresponding target data contains the observed hospitalization counts for these dates and locations.

```{r target_data}
target_data <- hubExamples::forecast_target_ts |>
  dplyr::filter(
    target_end_date %in% unique(forecast_data$target_end_date),
    location %in% unique(forecast_data$location),
    target == "wk inc flu hosp"
  ) |>
  # Rename columns to match the oracle output format
  rename(oracle_value = observation)

tbl_style(target_data)
```

We plot the point (median) forecasts and the observed values to visualize the difference in forecast errors.

```{r fig-example-median-lomo}
target_data |>
  mutate(model_id = "Observed") |>
  rename(value = oracle_value) |>
  rbind(
    forecast_data |>
      select("target_end_date", "target", "location", "value", "model_id")
  ) |>
  ggplot(aes(x = target_end_date)) +
  geom_point(aes(y = value, color = model_id, shape = model_id), size = 2) +
  geom_point(
    data = target_data,
    aes(y = oracle_value, color = "Observed", shape = "Observed"),
    size = 3
  ) +
  facet_wrap(~location,
    scales = "free_y",
    labeller = labeller(location = function(x) paste0("Location: ", x))
  ) +
  scale_x_date(
    breaks = target_data$target_end_date,
    date_labels = "%Y-%m-%d",
    expand = expansion(add = c(5, 5))
  ) +
  scale_color_manual(
    name = "model_id",
    values = c(
      "Flusight-baseline" = "#F8766D",
      "MOBS-GLEAM_FLUH" = "#00BA38",
      "PSI-DICE" = "#619CFF",
      "Observed" = "black"
    ),
    limits = c(
      "Flusight-baseline",
      "MOBS-GLEAM_FLUH",
      "PSI-DICE",
      "Observed"
    )
  ) +
  scale_shape_manual(
    name = "model_id",
    values = c(
      "Flusight-baseline" = 16,
      "MOBS-GLEAM_FLUH" = 17,
      "PSI-DICE" = 15,
      "Observed" = 18
    ),
    limits = c(
      "Flusight-baseline",
      "MOBS-GLEAM_FLUH",
      "PSI-DICE",
      "Observed"
    )
  ) +
  labs(
    y = "Weekly Hospitalization",
    x = "Date",
    title = "Forecasts of incident deaths generated on November 19, 2022"
  )
```

When comparing the ground truth data and model predictions, we can see that forecasts for December 10, 2022 show larger deviations from the observed values compared to those for November 26, 2022. Thus, as we expect, prediction errors increase at longer horizons due to greater uncertainty. 
Additionally, the forecasts for Massachusetts are relatively more accurate compared to those for Texas, which tend to have higher errors.


## Evaluation using LOMO algorithm

We quantify the contribution of each model within the ensemble using the `model_importance()` function. 
The following code evaluates the importance of each ensemble member in the simple mean ensemble using the LOMO algorithm. 

```{r lomo}
scores_lomo <- model_importance(
  forecast_data = forecast_data,
  oracle_output_data = target_data,
  ensemble_fun = "simple_ensemble",
  importance_algorithm = "lomo"
)

print(scores_lomo)
```

For models that missed forecasts for certain tasks, `NA` values were assigned in the importance column for those tasks.

Calling `summary()` shows that three models were used and four tasks were evaluated, along with a preview of the top-performing model for each task.

```{r}
summary(scores_lomo)
```

As indicated in the output, more details about the summary are available through the summary object's elements as follows.

```{r}
s <- summary(scores_lomo)
s$all_tasks
```


Each row represents a unique combination of task IDs, from which we verify that four different tasks were evaluated. 
```{r}
s$model_summary
```

We observe that 'Flusight-baseline' submitted forecasts for all four tasks (`n_NA` = 0), while 'MOBS-GLEAM_FLUH' and 'PSI-DICE' submitted forecasts for only three tasks due to one missing forecast (`n_NA` = 1). Each model's importance scores vary across tasks. 'Flusight-baseline' shows the largest range of scores that includes a negative minimum value and a positive maximum value, while 'MOBS-GLEAM_FLUH' and 'PSI-DICE' have scores that are all positive or all negative across the three tasks they submitted forecasts for.

```{r}
s$task_winners
```

Models with the highest importance scores for each task are identified in the `top_model` column with their importance score in the `max_score` column. 'PSI-DICE' is the best model for three out of the four tasks, while 'Flusight-baseline' is the best for the remaining task.

Visualization can be performed using functions from `ggplot2`. 
The following example shows a bar plot of importance scores across models and tasks, with panels faceted by combinations of task ID values.    

```{r}
ggplot(scores_lomo, aes(x = model_id, y = importance, fill = model_id)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.25) +
  facet_grid(
    cols = vars(target, horizon, location, target_end_date),
    scales = "free_x"
  ) +
  labs(
    x = "Model ID", y = "Importance Score",
    title = "Model Importance by Task"
  )
```


We aggregate the importance scores for each model by averaging across all tasks. `NA` values are removed during the averaging process by setting the `na_action` argument to `"drop"`.

```{r}
aggregate(scores_lomo, by = "model_id", na_action = "drop", fun = mean)
```

The results show that, overall, the model 'PSI-DICE' has the highest importance score, followed by 'Flusight-baseline' and 'MOBS-GLEAM_FLUH'.
That is, 'PSI-DICE' contributes the most to improving the ensemble's predictive performance, whereas 'MOBS-GLEAM_FLUH', which has a negative score, detracts from the ensemble’s performance.
The low importance score of 'MOBS-GLEAM_FLUH' is mainly due to a substantially larger prediction error for Texas on the target end date of December 10, 2022, compared to other models, while its missing forecast for Massachusetts for November 26, 2022, was not factored into the evaluation. This single large error significantly affected its contribution score.

Another approach to handling missing values is to use the `"worst"` option for `na_action`, which replaces missing values with the worst (i.e., minimum) score among the other models for the same task.

```{r}
aggregate(scores_lomo, by = "model_id", na_action = "worst", fun = mean)
```

It is also possible to impute the missing scores with intermediate values by assigning the average importance scores of other models in the same task. 
This strategy may offer a more balanced trade-off by mitigating the influence of the missing data without overly penalizing or overlooking them. 

```{r}
aggregate(scores_lomo, by = "model_id", na_action = "average", fun = mean)
```

## Evaluation using LASOMO algorithm

We now demonstrate the use of the LASOMO algorithm for evaluating model importance. 
Since we explored the difference of `na_action` options in the previous LOMO example [Section 6.2](#sec:example-lomo), we focus on options for `subset_wt`, which specifies how weights are assigned to subsets of models when calculating importance scores, with `na_action` fixed to `"drop"`.

The following code and corresponding outputs illustrate the evaluation using each weighting scheme.

```{r code-model_importance-summary-lasomo-equal}
scores_lasomo_eq <- model_importance(
  forecast_data = forecast_data,
  oracle_output_data = target_data,
  ensemble_fun = "simple_ensemble",
  importance_algorithm = "lasomo",
  subset_wt = "equal"
)
aggregate(scores_lasomo_eq, by = "model_id", na_action = "drop", fun = mean)
```

```{r code-model_importance-summary-lasomo-perm}
scores_lasomo_perm <- model_importance(
  forecast_data = forecast_data,
  oracle_output_data = target_data,
  ensemble_fun = "simple_ensemble",
  importance_algorithm = "lasomo",
  subset_wt = "perm_based"
)
aggregate(scores_lasomo_perm, by = "model_id", na_action = "drop", fun = mean)
```


In this example, there are only three models ($n = 3$), and the weights do not  differ significantly between the two weighting schemes. 
Therefore, the resulting outputs show little difference. 
However, in general, with a larger number of models, the two weighting schemes may yield quite different importance scores for each model, as discussed in [Comparison of weighting schemes in LASOMO](../articles/modelimportance-article.html#sec:comparison-weighting-schemes) section of the accompanying article.

Note that the computational time here is about 0.3 seconds for both LOMO and LASOMO algorithms.
However, this time can be increased substantially with a large number of models and tasks. 
See detailed discussions on execution time and computational feasibility in the 
[Computational complexity](../articles/modelimportance-article.html#sec:computational-complexity) section of the accompanying article.

An extensive application in more complex scenarios with a larger number of models can be found in @kim2026beyond.

## References
